Generation of TypeScript Declaration Files from JavaScript Code

### 2019-12-18 ecoop-2020/lipics-v2019-sample-article.tex
Add motivating example

\relatedversion{}%optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversion{A full version of the paper is available at \url{...}.}

@@

This work presents \lstinline[language={}]!dts-generate!, a tool that generates a declaration file for JavaScript libraries uploaded to the NPM registry. It executes the library, gathers run-time information from the code examples written by the developers and generates the declaration file based on it. 244 declaration files were generated and compared against DefinitelyTyped, the standard public repository for declaration files. 33 files out of 244 had no differences.
\end{abstract}

@@

\section{Motivating Examples}

@@

\subsection{TypeScript Declaration File Generation}
\paragraph*{Overwiew}
The actual generation of the declaration file is the next step in the pipeline after gathering the run-time information, as shown in \figref{fig:tsd_generation_method_block_diagram}. It is a lightweight, simple and fast application, which does not interact with the actual JavaScript module at run-time. Instead, it uses the JSON output file containing the run-time information and generates a TypeScript declaration file which is use ready to be used within a TypeScript project. The tool itself is written in TypeScript and runs within a Docker container in NodeJS.

@@

Templates of type \lstinline{module-class} are shown for modules \lstinline{flake-idgen}, \lstinline{route-parser} and \lstinline{timer-machine} in \figref{fig:experiments-results-module-class-flake-idgen}, \figref{fig:experiments-results-module-class-route-parser} and \figref{fig:experiments-results-module-class-timer-machine}, respectively. Properties of interfaces and class methods are correctly generated. Optional parameters are not detected, as it was not considered for the implementation. Finally, \lstinline{module} template is presented for \lstinline{is-uuid} module in \figref{fig:experiments-results-module-is-uuid}.

@@

\begin{figure}[tp]
    \centering
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export = Route;

declare namespace Route {
}
      \end{lstlisting}
      \caption{route-parser/index.d.ts - Generated}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
declare class Route {
    constructor(spec: string);
    match(pathname: string): { [i: string]: string } | false;

    reverse(params: { [i: string]: any }): string | false;
}

declare namespace Route { }
export = Route;       
      \end{lstlisting}
      \caption{route-parser/index.d.ts - DefinitelyTyped}
    \end{subfigure}

    \caption{\textbf{Module-class results | route-parser} - Constructor and methods are correctly identified. Specific signature of dictionary \lstinline{params} is not detected. Instead, a generic \lstinline{object} is inferred.}
    \label{fig:experiments-results-module-class-route-parser}
\end{figure}

>>>>

\relatedversion{}%optional, e.g. full version hosted on arXiv, HAL, or other repository/website
%\relatedversion{A full version of the paper is available at \url{...}.}

@@

This work presents \lstinline{dts-generate}, a tool that generates a declaration file for JavaScript libraries uploaded to the NPM registry. It executes the library, gathers run-time information from the code examples written by the developers and generates the declaration file based on it. 244 declaration files were generated and compared against DefinitelyTyped, the standard public repository for declaration files. 33 files out of 244 had no differences.
\end{abstract}

@@

\section{Motivating Example}
The NPM package \lstinline{route-parser} is a simple route parsing, matching, and reversing library for Javascript\footnote{https://www.npmjs.com/package/route-parser}. It has about 35000 weekly downloads and 221 NPM packages depend on it. If a developer is creating or extending a JavaScript library written in TypeScript that depends on \lstinline{route-parser}, the TypeScript compiler and IDEs will need a declaration file for that JS library in order to perform static checking and code completion, respectively. We use \lstinline{dts-generate} to automatically generate a TypeScript Declaration File for this library. The tool will download the npm package, run it, gather run-time information and generate a valid TypeScript Declaration File. \figref{fig:motivating-example-route-parser-vscode} shows the generated declaration file for package \lstinline{route-parser} using \lstinline{dts-generate}. The result is a valid declaration file that can be used in a TypeScript project. For example, it can be seen in \figref{fig:motivating-example-route-parser-vscode} that Visual Studio Code performs code completion properly. Finally, if \lstinline{route-parser} gets modified in a future, a new declaration file could be automatically generated using \lstinline{dts-generate}.

\begin{figure}[tp]
  \centering
  \begin{subfigure}{0.70\linewidth}
    \begin{lstlisting}[language=TypeScript]
export = Route;

declare namespace Route {
}
    \end{lstlisting}
  \end{subfigure}

  \hspace{10.\textwidth}

  \begin{subfigure}{1.\linewidth}
    \centering
    \includegraphics[width=0.7\linewidth]{motivating-example-route-parser-vscode.png}
  \end{subfigure}

  \caption{\textbf{Declaration file for \lstinline{route-parser} generated with \lstinline{dts-generate}} - Constructor and methods are correctly identified. Declaration file can be correctly used in Visual Studio Code\footnote{https://code.visualstudio.com}.}
  \label{fig:motivating-example-route-parser-vscode}
\end{figure}

@@

\subsection{TypeScript Declaration File Generation}
\paragraph*{Overview}
The actual generation of the declaration file is the next step in the pipeline after gathering the run-time information, as shown in \figref{fig:tsd_generation_method_block_diagram}. It is a lightweight, simple and fast application, which does not interact with the actual JavaScript module at run-time. Instead, it uses the JSON output file containing the run-time information and generates a TypeScript declaration file which is use ready to be used within a TypeScript project. The tool itself is written in TypeScript and runs within a Docker container in NodeJS.

@@

Templates of type \lstinline{module-class} are shown for modules \lstinline{flake-idgen}, \lstinline{route-parser} and \lstinline{timer-machine} in \figref{fig:experiments-results-module-class-flake-idgen} and \figref{fig:experiments-results-module-class-timer-machine}, respectively. Properties of interfaces and class methods are correctly generated. Optional parameters are not detected, as it was not considered for the implementation. Finally, \lstinline{module} template is presented for \lstinline{is-uuid} module in \figref{fig:experiments-results-module-is-uuid}.

@@
